---
title: "Explore SoCo Temperature"
author: "whalen"
date: "June 12, 2015"
output: html_document
---

```{r load data}
long_temp <- readRDS("Temperature/eof_pred_long.rds")# long format
# wide_temp <- readRDS("Temperature/eof_pred_wide.rds")
```

```{r examine/compare data structure, message=FALSE}
# str(long_temp)
# str(wide_temp)
# names(long_temp)
# names(wide_temp)
# summary(long_temp)
# summary(wide_temp)

library(plyr); library(dplyr)
long_temp <- select (long_temp, -posix_time)
long_temp$date <- as.Date(long_temp$chr_date, format = "%Y-%m-%d %H")
```


## Subset data to rainy season temperature variables for each sampling year

First, I added a variable to indicate the year of disease sampling the rainy season range of dates represents. The "rainy season" is defined here as November 1st - May 31st.

```{r create sample season indicator, message=FALSE}
library(dplyr)
rainy_season_temps <-  long_temp %>% mutate(sample_year = ifelse(
      date >= "2003-11-01" & date <= "2004-05-31", 2004, 
      ifelse(date >= "2004-11-01" & date <= "2005-05-31", 2005,
             ifelse(date >= "2005-11-01" & date <= "2006-05-31", 2006,
                    ifelse(date >= "2006-11-01" & date <= "2007-05-31", 2007,
                           ifelse(date >= "2007-11-01" & date <= "2008-05-31", 2008,
                                  ifelse(date >= "2008-11-01" & date <= "2009-05-31", 2009,
                                         ifelse(date >= "2009-11-01" & date <= "2010-05-31", 2010,
                                                ifelse(date >= "2010-11-01" & date <= "2011-05-31", 2011, ifelse(date >= "2011-11-01" & date <= "2012-05-31", 2012,
     ifelse(date >= "2012-11-01" & date <= "2013-05-31", 2013,                                                                                                      ifelse(date >= "2013-11-01" & date <= "2014-05-31", 2014, NA))))))))))))

rainy_season_temps <- filter(rainy_season_temps, is.na(sample_year)==F)
summary(rainy_season_temps)
```

Now I use the `sample_year` as a grouping variable to calculate the various temperature variables.

```{r calculate rainy season hour counts}
# hours within "optimal" temperature, 14-20 C
rs_select_temps <- rainy_season_temps %>% 
      filter(temp >= 14 & temp <= 20) %>% 
      group_by(plotid, sample_year) %>% 
      summarise(hrs_14_20 = length(temp))

# hours within "optimal" temperature, 10-20 C
rs_select_temps <- left_join(rs_select_temps, 
                             rainy_season_temps %>% 
                                   filter(temp >= 10 & temp <= 20) %>% 
                                   group_by(plotid, sample_year) %>%
                                   summarise(hrs_10_20 = length(temp)),
                             by = c("plotid", "sample_year"))

# hours at dessicating, slow/no growth temperatures, ≥ 25 C
rs_select_temps <- left_join(rs_select_temps, rainy_season_temps %>% 
                                  filter(temp >= 25) %>%
                                  group_by(plotid, sample_year) %>%
                                  summarise(hrs_abv25 = length(temp)),
                            by = c("plotid", "sample_year"))

# hours at/above dessicating, slow/no growth temperatures, ≥ 20 C
rs_select_temps <- left_join(rs_select_temps, rainy_season_temps %>% 
                                  filter(temp >= 20) %>%
                                  group_by(plotid, sample_year) %>%
                                  summarise(hrs_abv20 = length(temp)),
                            by = c("plotid", "sample_year"))

# hours below dessicating, slow/no growth temperatures, < 20 C
rs_select_temps <- left_join(rs_select_temps, rainy_season_temps %>% 
                                  filter(temp < 20) %>%
                                  group_by(plotid, sample_year) %>%
                                  summarise(hrs_blw20 = length(temp)),
                            by = c("plotid", "sample_year"))

# hours at cool slow growing temperatures, ≤ 10 C
rs_select_temps <- left_join(rs_select_temps, 
                                rainy_season_temps %>% 
                                      filter(temp <= 10) %>% 
                                      group_by(plotid, sample_year) %>%
                                      summarise(hrs_blw10 = length(temp)),
                                by = c("plotid", "sample_year"))

# hours at cool slow growing temperatures, ≤ 14 C
rs_select_temps <- left_join(rs_select_temps, 
                                rainy_season_temps %>% 
                                      filter(temp <= 14) %>% 
                                      group_by(plotid, sample_year) %>%
                                      summarise(hrs_blw14 = length(temp)),
                                by = c("plotid", "sample_year"))


rs_select_temps <- left_join(rs_select_temps, rainy_season_temps %>% 
                                  group_by(plotid, sample_year, date) %>% 
      summarise(daily_tmax = max(temp), daily_tmin = min(temp)) %>% 
      group_by(plotid, sample_year) %>% 
      summarise(avg_tmax = mean(daily_tmax), avg_tmin = mean(daily_tmin)),
      by = c("plotid", "sample_year"))


summary(rs_select_temps)

write.csv(rs_select_temps, "~/GitHub/superspreaders/analysis/data/rs_select_temps.csv", row.names = F)
```


```{r plotting}
library(ggplot2)
library(scales)
# Subset to fewer plots
dt <- long_temp %>%
      select(year, temp, plotid, date) %>%
      group_by(as.factor(year)) %>%
      filter(plotid == "ANN01" | plotid == "SUGAR23" | plotid == "FOP04")

qplot(x = year, y = avg_temp_yr, 
      data = dt %>% 
            group_by(year, plotid) %>%
            select(year, plotid, temp) %>% 
            summarise(avg_temp_yr = mean(temp)),
      facets = plotid ~ ., color = plotid, geom = c("point", "line")
            )

tempXyear <- ggplot(long_temp %>% filter(year != 2003), aes(date, temp))

tempXyear + 
      geom_line() +
      scale_x_date(date_breaks = "1 year", 
                   labels = date_format("%m/%y"),
                   date_minor_breaks = "6 month") +
      xlab("Date (month/year)") +
      ylab("Temperature (degrees Celsius)")
```

## Subset data to summer season temperature variables for each sampling year

First, I added a variable to indicate the year of disease sampling the summer season range of dates represents. The "summer season" is defined here as June 1st - September 30th. Summer temperatures are most likely to influence the oversummering of the pathogen on leaves due to increasing abscision of infected leaves under hotter conditions. The infected leaves that survive contribute to the inoculum capacity the following season, because the surviving infections enable pathogen reproduction to progress more quickly when wet conditions return. Based on this I'm calculating temperature aggregates for summers 2003 - 2013, which can be used in models as variables predicting infection levels the following year. E.g., the 2003 temperatures would be used as a predictor for the 2004 symptomatic leaf count.

I'm going to label the dates according to the sample year so that they match up with the ecological data more easily.

```{r create sample season indicator, message=FALSE}
library(dplyr)
summer_season_temps <- long_temp %>% 
      mutate(sample_year = ifelse(
            date >= "2003-06-01" & date <= "2003-09-30", 2004, 
            ifelse(date >= "2004-06-01" & date <= "2004-09-30", 2005,
              ifelse(date >= "2005-06-01" & date <= "2005-09-30", 2006,
                ifelse(date >= "2006-06-01" & date <= "2006-09-30", 2007,
                  ifelse(date >= "2007-06-01" & date <= "2007-09-30", 2008,
                    ifelse(date >= "2008-06-01" & date <= "2008-09-30", 2009,
                      ifelse(date >= "2009-06-01" & date <= "2009-09-30", 2010,
        ifelse(date >= "2010-06-01" & date <= "2010-09-30", 2011, 
          ifelse(date >= "2011-06-01" & date <= "2011-09-30", 2012,
                 ifelse(date >= "2012-06-01" & date <= "2012-09-30", 2013,
            ifelse(date >= "2013-06-01" & date <= "2013-09-30", 2014, NA)
)))))))))))

summer_season_temps <- (filter(summer_season_temps, is.na(sample_year)==F))
summary(summer_season_temps)
```


```{r calculate summer temp hour counts}
# hours at 'extreme' desiccating, slow/no growth temperatures, >= 25 C (~77 F)
# Most likely to enhance leaf abscision, especially of infected

# summer_select_temps <- summer_season_temps %>% 
#       filter(temp >= 30) %>%
#       group_by(plotid, sample_year) %>%
#       summarise(hrs_abv30 = length(temp))

summer_select_temps <- summer_season_temps %>% 
      filter(temp >= 25) %>%
      group_by(plotid, sample_year) %>%
      summarise(hrs_abv25 = length(temp))

# hours at desiccating, slow/no growth temperatures, >= 20 C (~68 F)
summer_select_temps <- left_join(summer_select_temps, summer_season_temps %>% 
                                  filter(temp >= 20) %>%
                                  group_by(plotid, sample_year) %>%
                                  summarise(hrs_abv20 = length(temp)),
                            by = c("plotid", "sample_year"))

# hours at cool, 'persisting' temperatures, < 20 C (~68 F)
summer_select_temps <- left_join(
      summer_select_temps, summer_season_temps %>%
            filter(temp < 20) %>% 
            group_by(plotid, sample_year) %>%
            summarise(hrs_blw20 = length(temp)),
      by = c("plotid", "sample_year"))

summer_select_temps <- left_join(
      summer_select_temps, summer_season_temps %>% 
            group_by(plotid, sample_year, date) %>% 
            summarise(daily_tmax = max(temp), daily_tmin = min(temp)) %>%
            group_by(plotid, sample_year) %>% 
            summarise(avg_tmax = mean(daily_tmax), 
                      avg_tmin = mean(daily_tmin)),
      by = c("plotid", "sample_year"))

summary(summer_select_temps)

write.csv(summer_select_temps, "~/GitHub/superspreaders/analysis/data/summer_select_temps.csv", row.names = F)
```

